---
title: "The Economy in Charts"
author: "Stephen Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document2:
    number_sections: true
    keep_tex: true
    toc: false
  word_document:
    toc: false
header-includes: 
- \usepackage{graphicx}
- \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message=FALSE,
    warning=FALSE,
    fig.align='center',
    fig.pos = 'H',
    fig.height=2, 
    fig.width=7
)
```

```{r plot-tools}
#-------------------------------- SECTION SETUP --------------------------------
library(tidyverse)
library(gridExtra)
library(fredr)
library(here)

source(here("utils", "config.R"))
source(here("utils", "pltlib.R"))
source(here("utils", "theme_wsj.R"))
source(here("utils", "label_wsj.R"))

fredr_set_key(FRED_KEY)

# STORAGE 
start_date <- as.Date("1995-01-01")
unit_change <- "pc1"
unit_level <- "lin"
label_percent <- "Percent"
label_dollars <- "USD"
recession <- "USRECM"
recession_dates <- get_recessions(start_date = start_date)
```

\newpage
# Monetary Policy 
```{r mp}
#-------------------------------- SECTION SETUP --------------------------------
ten_two_year_spread <- "T10Y2Y"
ten_year_level      <- "DGS10"
two_year_level      <- "DGS2"
effective_ffr       <- "DFF"
```

## Interest Rates
The federal funds rate is the interest rate at which depository institutions trade federal funds (balances held at Federal Reserve Banks) with each other overnight. When a depository institution has surplus balances in its reserve account, it lends to other banks in need of larger balances. In simpler terms, a bank with excess cash, which is often referred to as liquidity, will lend to another bank that needs to quickly raise liquidity. The rate that the borrowing institution pays to the lending institution is determined between the two banks; the weighted average rate for all of these types of negotiations is called the effective federal funds rate. The effective federal funds rate is essentially determined by the market but is influenced by the Federal Reserve through open market operations to reach the federal funds rate target.

Treasury yields are of 2 and 10-Year constant maturity rates. 
```{r tenlevel}
d <- c(effective_ffr, two_year_level, ten_year_level)
labels <- c("FFR", "2Y", "10Y")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels = labels)
```

## Ten Year Minus Two Year
10-Year Treasury Constant Maturity Minus 2-Year Treasury Constant Maturity
```{r tentwo}
plt(ten_two_year_spread, start_date, unit_level, recession_dates)
```

\newpage
# Inflation 
```{r inflation}
#-------------------------------- SECTION SETUP --------------------------------
cpi                 <- "FPCPITOTLZGUSA"
pce                 <- "PCETRIM12M159SFRBDAL"
five_yr_breakeven   <- "T5YIE"
fivey_fivey         <- "T5YIFR"
```

## CPI and PCE
Inflation as measured by the consumer price index reflects the annual percentage change in the cost to the average consumer of acquiring a basket of goods and services that may be fixed or changed at specified intervals, such as yearly. The Laspeyres formula is generally used.

The Trimmed Mean PCE inflation rate produced by the Federal Reserve Bank of Dallas is an alternative measure of core inflation in the price index for personal consumption expenditures (PCE). The data series is calculated by the Dallas Fed, using data from the Bureau of Economic Analysis (BEA). Calculating the trimmed mean PCE inflation rate for a given month involves looking at the price changes for each of the individual components of personal consumption expenditures. The individual price changes are sorted in ascending order from “fell the most” to “rose the most,” and a certain fraction of the most extreme observations at both ends of the spectrum are thrown out or trimmed. The inflation rate is then calculated as a weighted average of the remaining components. The trimmed mean inflation rate is a proxy for true core PCE inflation rate. The resulting inflation measure has been shown to outperform the more conventional “excluding food and energy” measure as a gauge of core inflation.
```{r pce}
d <- c(cpi, pce)
labels <- c("CPI", "PCE")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels)
```

## Inflation Expectations
The breakeven inflation rate represents a measure of expected inflation derived from 5-Year Treasury Constant Maturity Securities and 5-Year Treasury Inflation-Indexed Constant Maturity Securities. The latest value implies what market participants expect inflation to be in the next 5 years, on average. 

The Five Year - Five Year Forward (5Y5Y) is a measure of expected inflation (on average) over the five-year period that begins five years from today.

```{r futureinflation}
d <- c(five_yr_breakeven, fivey_fivey)
labels <- c("5Y Breakeven", "5Y5Y")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels)
```

\newpage
# Financial Markets 
```{r finmarkets}
#-------------------------------- SECTION SETUP --------------------------------
sp500               <- "SP500"
nasdaq              <- "NASDAQCOM"
micro_cap           <- "WILLMICROCAP"
small_cap           <- "WILLSMLCAPPR"
mid_cap             <- "WILLMIDCAPPR"
```

## S&P 500 Index 
The S&P 500 is regarded as a gauge of the large cap U.S. equities market. The index includes 500 leading companies in leading industries of the U.S. economy, which are publicly held on either the NYSE or NASDAQ, and covers 75% of U.S. equities. Since this is a price index and not a total return index, the S&P 500 index here does not contain dividends.
```{r sp500}
plt(sp500, start_date, unit_level, recession_dates, "Index",
  prefix = "", suffix = ""
)
```

## NASDAQ
The NASDAQ Composite Index is a market capitalization weighted index with more than 3000 common equities listed on the NASDAQ Stock Market. The types of securities in the index include American depositary receipts (ADRs), common stocks, real estate investment trusts (REITs), and tracking stocks. The index includes all NASDAQ listed stocks that are not derivatives, preferred shares, funds, exchange-traded funds (ETFs) or debentures.
```{r nasdaq}
plt(nasdaq, start_date, unit_level, recession_dates, "Index",
  prefix = "", suffix = ""
)
```

## Wilshire Index - Mid Cap
```{r midcap}
plt(mid_cap, start_date, unit_level, recession_dates, "Index", 
    prefix = "", suffix = ""
)
```

## Wilshire Index - Small Cap
```{r smallcap}
plt(small_cap, start_date, unit_level, recession_dates, "Index", 
    prefix = "", suffix = ""
)
```

## Wilshire Index - Micro Cap
```{r microcap}
plt(micro_cap, start_date, unit_level, recession_dates, "Index", 
    prefix = "", suffix = ""
)
```

\newpage
# Corporate Banking 
```{r corpbank}
#-------------------------------- SECTION SETUP --------------------------------
# corporate loan info 
ci_loans            <- "TOTCI"
ci_loans_small      <- "ACILOB"

# deliquiency info 
ci_delinq_all       <- "DRBLACBS"
ci_delinq_top       <- "DRBLT100S"
ci_delinq_bottom    <- "DRBLOBS"

# charge off rates 
nci_chargeoff        <- "NCOALLCIACB"

# sloos
ci_tight_stds_lg     <- "DRTSCILM"
ci_tight_stds_sm     <- "DRTSCIS"
ci_demand_lg         <- "DRSDCILM"
ci_demand_sm         <- "DRSDCIS"
```

## C&I Loans 
The percent change from a year ago. 

```{r ci}
d <- c(ci_loans, ci_loans_small)
labels <- c("All Banks", "Small Banks")
names(labels) <- d
plt(d, start_date, unit_change, recession_dates, labels=labels)
```

## C&I Delinquency Rate

```{r cidelinq}
d <- c(ci_delinq_all, ci_delinq_top, ci_delinq_bottom)
labels <- c("All Banks", "Top 100 Banks", "Excluding Top 100")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels = labels)
```

## C&I Charge-Offs 
Charge-offs are the value of loans and leases removed from the books and charged against loss reserves. This includes all commercial banks. 

```{r chargeoffs}
plt(nci_chargeoff, start_date, unit_level, recession_dates, 
    prefix = "$ ", suffix = "M"
)
```

## SLOOS Reporting Tighter Standards - C&I Loans
This is the net percentage of domestic banks reporting tightening standards for commercial and industrial loans. "Large Firms" refers to loans made to large and middle-market firms. 
```{r sloosstds}
d <- c(ci_tight_stds_lg, ci_tight_stds_sm)
labels <- c("Large Firms", "Small Firms")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels)
```

## SLOOS Reporting Increased Demand - C&I Loans
This is the net percentage of domestic banks reporting stronger demand for commercial and industrial loans. "Large Firms" refers to loans made to large and middle-market firms.
```{r sloosdemand}
d <- c(ci_demand_lg, ci_demand_sm)
labels <- c("Large Firms", "Small Firms")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels)
```

\newpage
# Corporate Bonds 
```{r corpbond}
#-------------------------------- SECTION SETUP --------------------------------
aaa                  <- "BAMLC0A1CAAA"
bbb                  <- "BAMLC0A4CBBB"
bb                   <- "BAMLH0A1HYBB"
b                    <- "BAMLH0A2HYB"
ccc                  <- "BAMLH0A3HYC"
corp_1to3_yr         <- "BAMLC1A0C13Y"
corp_7to10_yr        <- "BAMLC4A0C710Y"
```

## Option Adjusted Spreads - Investment Grade
The ICE BofAML Option-Adjusted Spreads (OASs) are the calculated spreads between a computed OAS index of all bonds in a given rating category and a spot Treasury curve. An OAS index is constructed using each constituent bond‚ OAS, weighted by market capitalization.
```{r igspreads}
d <- c(aaa, bbb)
labels <- c("AAA", "BBB")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels)
```

## Option Adjusted Spreads - High Yield
```{r junkspreads}
d <- c(bb, b, ccc)
labels <- c("BB", "B", "CCC")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels)
```

## Forward Looking Option Adjusted Spreads
The US Corporate 1-3 Year OAS is a subset of the ICE BofAML US Corporate Master OAS, BAMLC0A0CM. This subset includes all securities with a remaining term to maturity of less than 3 years. When the last calendar day of the month takes place on the weekend, weekend observations will occur as a result of month ending accrued interest adjustments. Similar arguement for the 7-10 Year. 
```{r forwardspreads}
d <- c(corp_1to3_yr, corp_7to10_yr)
labels <- c("1-3 Years", "7-10 Years")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels, 
    prefix = "", suffix = ""
)
```

\newpage
# Housing 
```{r housing}
#-------------------------------- SECTION SETUP --------------------------------
national             <- "CSUSHPISA"
atl                  <- "ATXRSA"
sf                   <- "SFXRSA"
chi                  <- "CHXRSA"
seattle              <- "SEXRSA"
dallas               <- "DAXRSA"
detroit              <- "DEXRSA"
cleveland            <- "CEXRSA"

# mortgage rates 
thirty_yr_fixed      <- "MORTGAGE30US"
fifteen_yr_fixed     <- "MORTGAGE15US"
```

## Mortgage Rates 
```{r mortgage}
d <- c(fifteen_yr_fixed, thirty_yr_fixed)
labels <- c("15Y", "30Y")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels)
```

## S&P/Case-Shiller Home Price Index
```{r hpi}
d <- c(national, atl, sf, chi, detroit)
labels <- c("National", "Atlanta", "San Francisco", "Chicago", "Detroit")
names(labels) <- d
plt(d, start_date, unit_level, recession_dates, labels=labels, "Index",  
    prefix = "", suffix = ""
)
```

\newpage
# Labor Markets
```{r labor}
#-------------------------------- SECTION SETUP --------------------------------
prof_services_emp    <- "USPBS"  # want pct change from year ago 
prime_age_emp        <- "LREM25TTUSM156S" # level and pct change
prime_age_male_emp   <- "LREM25MAUSM156S" # level and pct change 
```

## Prime Age Employment - Aged 25-54
Percent change in employment from a year ago. 

```{r primeempl}
d <- c(prime_age_emp, prime_age_male_emp)
labels <- c("All", "Males")
names(labels) <- d
plt(d, start_date, unit_change, recession_dates, labels=labels)
```

## Professional Services Employment
Percent change in employment from a year ago. 

```{r profserivesemp}
plt(prof_services_emp, start_date, unit_change, recession_dates)
```
